---
output: github_document
---

```{r}
#Load libraries 
library(dplyr)
library(tidyr)
library(ggplot2)
theme_set(theme_bw())

#Load in ideal data 
load("tidy_df.Rdata")
tidy_df_ideal = tidy_df
#remove col to make same # cols 
tidy_df_ideal = subset(tidy_df_ideal, select = -c(seeds_per_tree))

#Realistic data
load("tidy_df_realistic.Rdata")
tidy_df_realistic = tidy_df

#Create new columns to identify scenarios 
tidy_df_ideal$samp_type = "ideal"
tidy_df_realistic$samp_type = "realistic"

#Combine dataframes
combined_ideal_real = rbind(tidy_df_ideal, tidy_df_realistic)
```

```{r}
str(combined_ideal_real)
#Convert to factors for ANOVA
combined_ideal_real$donor_type = factor(combined_ideal_real$donor_type)
combined_ideal_real$maternal_trees = factor(combined_ideal_real$maternal_trees)
combined_ideal_real$samp_type = factor(combined_ideal_real$samp_type)
combined_ideal_real$prop_capt = as.numeric(combined_ideal_real$prop_capt)
```

```{r}
#defining more descriptive labels for the facets 
mat_tree_labs = c("2 maternal trees", "5 maternal trees", "10 maternal trees", "25 maternal trees", "50 maternal trees", "100 maternal trees")
names(mat_tree_labs) = c("2", "5", "10", "25", "50", "100")

#Plotting the data
combined_ideal_real %>%
    filter(total_seeds==200) %>%
    filter(as.numeric(maternal_trees)>1) %>%
    ggplot(aes(x=donor_type, y=as.numeric(prop_capt), color=samp_type)) +
    geom_boxplot() +
    facet_wrap(vars(as.numeric(maternal_trees))) +
    theme(strip.background = element_rect(color="black", fill="#F2F2F2", linetype="solid")) +
    ylim(0,1) + 
    ggtitle("Genetic diversity capture for scenarios with 200 total seeds sampled") +
    xlab("Donor type") +
    ylab("Proportion of alleles captured") +
    labs(color = "Sampling Type") +
    scale_x_discrete(labels=c("Eligible","Same","Skewed")) 
```

Compute ANOVA:
Factors: maternal trees sampled, donor type, and sample type? 
Response variable: proportion of alleles captured 
Interactions between all factors because they may all interact to create a synergistic effect 
```{r}
anova_result = aov(prop_capt ~ maternal_trees * donor_type * samp_type, data = combined_ideal_real)
summary(anova_result)
TukeyHSD(anova_result)
```

Checking assumptions
```{r}
#Residuals vs fits plot 
plot(anova_result, 1)
library(car)
#Levene test to check homogeneity of variances 
leveneTest(as.numeric(prop_capt) ~ maternal_trees * donor_type * samp_type, data = combined_ideal_real)
#Normality of residuals 
plot(anova_result, 2)
#Shapiro Wilke normality test
aov_residuals = residuals(object = anova_result)
shapiro.test(x = sample(aov_residuals, 5000))
```
Here we can see that the assumptions of the ANOVA are violated, so if proceeding, need to run equivalent non-parametric analyses. 
Also important to note that the study design is unbalanced, meaning that there are different numbers of subject in each group.

T-tests to compare equivalent scenarios between ideal and realistic: 

```{r}
p_vals_same = c()
ideal_avg = c()
realistic_avg = c()
mat_trees = c(2, 5, 10, 25, 50, 100)
for(i in 1:length(mat_trees)) {
    subset = combined_ideal_real %>% filter(total_seeds == 200) %>% filter(maternal_trees == mat_trees[i]) %>% filter(donor_type == 'all_same')
    ideal_avg[i] = mean((subset %>% filter(samp_type == "ideal"))$prop_capt)
    realistic_avg[i] = mean((subset %>% filter(samp_type == "realistic"))$prop_capt)
    p_vals_same[i] = t.test(subset$prop_capt ~ subset$samp_type)$p.value   
}
same_table = data.frame(donor_type = rep("same", 6), maternal_trees_samp = mat_trees, tot_seeds_sampled = rep(200, 6), prop_capt_ideal = ideal_avg, prop_capt_realistic = realistic_avg, p_value = p_vals_same)
same_table

p_vals_skewed = c()
mat_trees = c(2, 5, 10, 25, 50, 100)
for(i in 1:length(mat_trees)) {
    subset = combined_ideal_real %>% filter(total_seeds == 200) %>% filter(maternal_trees == mat_trees[i]) %>% filter(donor_type == 'skewed')
    ideal_avg[i] = mean((subset %>% filter(samp_type == "ideal"))$prop_capt)
    realistic_avg[i] = mean((subset %>% filter(samp_type == "realistic"))$prop_capt)
    p_vals_skewed[i] = t.test(subset$prop_capt ~ subset$samp_type)$p.value   
}
skewed_table = data.frame(donor_type = rep("skewed", 6), maternal_trees_samp = mat_trees, tot_seeds_sampled = rep(200, 6), prop_capt_ideal = ideal_avg, prop_capt_realistic = realistic_avg, p_value = p_vals_skewed)
skewed_table

p_vals_eligible = c()
mat_trees = c(2, 5, 10, 25, 50, 100)
for(i in 1:length(mat_trees)) {
    subset = combined_ideal_real %>% filter(total_seeds == 200) %>% filter(maternal_trees == mat_trees[i]) %>% filter(donor_type == 'all_eligible')
    ideal_avg[i] = mean((subset %>% filter(samp_type == "ideal"))$prop_capt)
    realistic_avg[i] = mean((subset %>% filter(samp_type == "realistic"))$prop_capt)
    p_vals_eligible[i] = t.test(subset$prop_capt ~ subset$samp_type)$p.value   
}
eligible_table = data.frame(donor_type = rep("eligible", 6), maternal_trees_samp = mat_trees, tot_seeds_sampled = rep(200, 6), prop_capt_ideal = ideal_avg, prop_capt_realistic = realistic_avg, p_value = p_vals_eligible)
skewed_table

p_values_combined = rbind(same_table, skewed_table, eligible_table)
```


